---
title: 'Part 2: Chi-squared test with lizzard tails'
author: "Laurel Wee"
date: "12/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(broom)
```
```{r}
lizards <- read_csv(here("data", "lizards.csv"))
```

Get counts of lizards by species

```{r}
lizards %>% 
  count(spp) %>% 
  arrange(-n)
```
## Make a subset

- UTST lizards
- Sites: IBPE (enclosure grassland), RABB (rabbit mesquite site), "EAST" (east tarbush site)
- Tail condition is recorded (W = whole, B = broken)

```{r}
utst <- lizards %>% 
  filter(spp == "UTST") %>% 
  filter(site %in% c("IBPE", "RABB", "EAST")) %>% 
  filter(tail %in% c("B","W"))

```

## Find counts and proportions of lizards with broken or whole tails at each site
```{r}
utst %>% 
  count(site,tail)

```
```{r}
tail_counts <- utst %>% 
  janitor::tabyl(site, tail)

tail_proportions <- tail_counts %>% 
  janitor::adorn_percentages() %>% # Default calculates proportions by rows
  janitor::adorn_pct_formatting(digits = 2) %>% 
  janitor::adorn_ns()
```

## Convert this to a contingency table

Use `column_to_rowname()`

```{r}
lizard_ct <- tail_counts %>% 
  column_to_rownames(var = "site")
```

## Chi-squared

```{r}
lizard_x2 <- chisq.test(lizard_ct)
lizard_x2

lizard_tidy <- broom::tidy(lizard_x2)
```

There is a significant association (i.e. non-independence) between site and lizard tail breakage ($\chi$^2^(`r lizard_tidy$parameter`) = `r lizard_tidy$statistic`, *p* = `r lizard_tidy$p.value`).



